\textbf{CSDS 491: Assignment 2}

### 2021-03-09 A2/article.tex
✨ Add answers for A2 Q5.

\section*{Q5}

>>>>

\section*{Q5. Car Troubles}

\subsection*{5.1.}

\begin{align*}
    p(f=\text{empty} \mid s=\text{no}) &= \frac{p(s=\text{no}, f=\text{empty})}{p(s=\text{no})}\\
    &= \frac{\sum_t\sum_b p(s=\text{no}, f=\text{empty}, \ t, b)}{\sum_f\sum_t\sum_b p(s=\text{no},f,t,b)}\\
    &= \frac{\sum_t\sum_b p(s=\text{no} \mid f=\text{empty},t)p(f, t)p(b)}{\sum_f\sum_t\sum_b p(s=\text{no} \mid f,t)p(f, t)p(b)}\\
    &= \frac{\sum_t\sum_b p(s=\text{no} \mid f=\text{empty},t) \ p(f)p(t \mid b)p(b)}{\sum_f\sum_t\sum_b p(s=\text{no} \mid f,t) \ p(f)p(t \mid b)p(b)}\\
    &= \frac{0.0461715}{0.101756} \approx 0.4537
\end{align*}

\subsection*{5.2.}

The following result generated by \ilc{code/q5.py} confirms my calculation of the pervious question.

\begin{lstlisting}
+-----------------+-------------+
| Fuel            |   phi(Fuel) |
+=================+=============+
| Fuel(not empty) |      0.5463 |
+-----------------+-------------+
| Fuel(empty)     |      0.4537 |
+-----------------+-------------+
\end{lstlisting}

\subsection*{5.3.}

Results within this question is generated by \ilc{code/q5.py}.

\subsubsection*{Scenario 1: No fuel, given Turns Over first.}

Given that \ilc{Turns Over = no} and known that \ilc{Starts = no}, we may query on the probability of \ilc{Battery} and \ilc{Fuel}:

\begin{lstlisting}
+---------------+----------------+          +-----------------+-------------+
| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
+===============+================+          +=================+=============+
| Battery(good) |         0.6000 |          | Fuel(not empty) |      0.9505 |
+---------------+----------------+          +-----------------+-------------+
| Battery(bad)  |         0.4000 |          | Fuel(empty)     |      0.0495 |
+---------------+----------------+         +-----------------+-------------+
\end{lstlisting}

So the problem is mostly likely caused by the car having no fuel. By asking our friend to provide additional information of \ilc{Gauge = not empty}, we may again query on the probability of \ilc{Battery} and \ilc{Fuel}:
\begin{lstlisting}
+---------------+----------------+          +-----------------+-------------+
| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
+===============+================+          +=================+=============+
| Battery(good) |         0.6156 |          | Fuel(not empty) |      0.9988 |
+---------------+----------------+          +-----------------+-------------+
| Battery(bad)  |         0.3844 |          | Fuel(empty)     |      0.0012 |
+---------------+----------------+          +-----------------+-------------+
\end{lstlisting}

We may notice that our confidence of the car having no fuel has further improved, and we have therefore found the true cause of the problem.

\subsubsection*{Scenario 2: Bad Battery, given Gauge first.}

Similar to the above setup but this time we were given that \ilc{Gauge = empty} first, and again known that \ilc{Starts = no}; we may query on the probability of \ilc{Battery} and \ilc{Fuel}:

\begin{lstlisting}
+---------------+----------------+          +-----------------+-------------+
| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
+===============+================+          +=================+=============+
| Battery(good) |         0.9410 |          | Fuel(not empty) |      0.0694 |
+---------------+----------------+          +-----------------+-------------+
| Battery(bad)  |         0.0590 |          | Fuel(empty)     |      0.9306 |
+---------------+----------------+          +-----------------+-------------+
\end{lstlisting}

We may tell the leading potential cause of the problem might be \ilc{Fuel = empty}. Now we may ask our friend to collect information on \ilc{Turns Over} where we were given that \ilc{Turns Over = no}, we may again query on the probability of \ilc{Battery} and \ilc{Fuel}:

\begin{lstlisting}
+---------------+----------------+          +-----------------+-------------+
| Battery       |   phi(Battery) |          | Fuel            |   phi(Fuel) |
+===============+================+          +=================+=============+
| Battery(good) |         0.4726 |          | Fuel(not empty) |      0.5567 |
+---------------+----------------+          +-----------------+-------------+
| Battery(bad)  |         0.5274 |          | Fuel(empty)     |      0.4433 |
+---------------+----------------+          +-----------------+-------------+
\end{lstlisting}

We observed that the probability of fuel being empty has dramatically decresed, and now the leading potential cause for the problem become \ilc{Battery = bad} -- and that's the true cause of the problem.
